---
title: "CARET_Lab"
output:
  pdf_document:
    latex_engine: xelatex
date: "27/11/2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Package loading
```{r}
library(caret)
```

Load Data
```{r}
# attach the iris dataset to the environment
data(iris)
# rename the dataset
dataset <- iris
```

Task1: Create a Validation/Training Dataset
You need to split the loaded dataset into two, 80% of which we will use to train our models and 20% that we will hold back as a validation dataset.
Hint: use createDataPartition function
```{r}

# Create the training and test datasets
set.seed(100)

# Step 1: Get row numbers for the training data
trainRowNumbers <- createDataPartition(dataset$Species, p=0.8, list=FALSE)

# Step 2: Create the training  dataset
trainData <- dataset[trainRowNumbers,]

# Step 3: Create the test dataset
testData <- dataset[-trainRowNumbers,]

```

Task2: Summarize Dataset
Use skimr library to summarize the dataset
```{r}

library(skimr)
skimmed <- skim_to_wide(trainData)

skimmed

```

Task3: split input and output
 It is the time to seperate the input attributes and  the output attributes. call the inputs attributes x and the output attribute (or class) y.
```{r}

# Store X and Y for later use.
x = trainData[, 1:4]
y = trainData[, 5]

```

Task4: Train Control for Validation Test

We will use 10-fold crossvalidation to estimate accuracy.
```{r}
# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

Task5: Model Training
Train 5 different algorithms using 'train' function:

- Linear Discriminant Analysis (LDA).
- Classification and Regression Trees (CART).
- k-Nearest Neighbors (kNN).
- Support Vector Machines (SVM) with a linear kernel.
- Random Forest (RF).

# Linear Discriminant Analysis (LDA).
```{r}

# Set the seed for reproducibility
set.seed(100)

# Train the model using LDA
model_LDA = train(Species ~ ., data=trainData, method = 'lda',  trControl = control, metric = metric)

model_LDA

```
# Classification and Regression Trees (CART).
```{r}

# Set the seed for reproducibility
set.seed(100)

# Train the model using CART
model_CART = train(Species ~ ., data=trainData, method = 'rpart',  trControl = control, metric = metric)

model_CART

```
# k-Nearest Neighbors (kNN).
```{r}

# Set the seed for reproducibility
set.seed(100)

# Train the model using kNN
model_KNN = train(Species ~ ., data=trainData, method = 'knn',  trControl = control, metric = metric)

model_KNN

```
# Support Vector Machines (SVM) with a linear kernel.
```{r}

# Set the seed for reproducibility
set.seed(100)

# Train the model using SVM
model_SVM = train(Species ~ ., data=trainData, method = 'svmRadial',  trControl = control, metric = metric)

model_SVM

```
# Random Forest (RF).
```{r}

# Set the seed for reproducibility
set.seed(100)

# Random Forest (RF)
model_RF = train(Species ~ ., data=trainData, method = 'rf',  trControl = control, metric = metric)

model_RF

```
Task6: Select the Best Model
We now have 5 models and accuracy estimations for each. We need to compare the models to each other and select the most accurate.
Use resamples function to complete this task

```{r}

#Compare all 5 using resample

models_compare <- resamples(list(LDA = model_LDA, CART = model_CART, KNN = model_KNN, SVMLinear = model_SVM, RF = model_RF))

summary(models_compare)

```
What was the most accurate model?
The most accurate model was the KNN model with 98.33% accuracy.


Task7: Make Prediction (Confusion Matrix)
Now we want to get an idea of the accuracy of the best model on our validation set. Use 'predict' and confusionMatrix functions to complete this task.

# LDA Confusion Matrix & Prediction
```{r}
predicted <- predict(model_LDA, testData)
confusionMatrix(reference = testData$Species, data = predicted)
```

# CART Confusion Matrix & Prediction
```{r}
predicted <- predict(model_CART, testData)
confusionMatrix(reference = testData$Species, data = predicted)
```

# KNN Confusion Matrix & Prediction
```{r}
predicted <- predict(model_KNN, testData)
confusionMatrix(reference = testData$Species, data = predicted)
```

# SVM Confusion Matrix & Prediction
```{r}
predicted <- predict(model_SVM, testData)
confusionMatrix(reference = testData$Species, data = predicted)
```

# RF Confusion Matrix & Prediction
```{r}
predicted <- predict(model_RF, testData)
confusionMatrix(reference = testData$Species, data = predicted)
```

